---
title: "Drawing M-graph assumptions"
author: "Janick Weberpals"
date: "Last compiled `r Sys.time()`"
output:
  html_document:
    toc: yes
    toc_depth: 2
    toc_float: yes
    number_sections: yes
    code_folding: show
    classoption: landscape
---

```{r, setup}
suppressMessages(library(dagitty))
suppressMessages(library(ggdag))
suppressMessages(library(ggplot2))
```


# Glossary

| Symbol | Meaning                                                     |
|--------|-------------------------------------------------------------|
| E      | Exposure/treatment                                          |
| Y      | Outcome                                                     |
| C      | Confounder of interest                                      |
| C_obs  | Observed portion of C                                       |
| M      | Missingness of C (M=0 fully observed and M=1 fully missing) |
| c1     | Covariates associated with outcome and missingness          |
| c0     | Auxiliary covariates                                        |
| U      | Unmeasured covariate/confounder                             |

# General coordinates
```{r}
# coordinates for mcar, mar, mnar_vaue
coords <- list(
  x = c(c1 = -1, E = 0, Y = 1, C = -1, C_obs = -1, M = -1),
  y = c(c1 =  2, E = 1, Y = 1, C =  0, C_obs = -1, M = -2)
  )
```


\newpage

# MCAR
```{r}
#  using more R-like syntax to create the same DAG
mcar <- ggdag::dagify(
  Y ~ E,
  Y ~ c1 + C,
  E ~ c1 + C,
  C ~ c1,
  C_obs ~ M + C,
  exposure = "E",
  outcome = "Y",
  coords = coords)  %>% 
  ggdag::tidy_dagitty() %>% 
  ggdag::ggdag_classic() +
  ggdag::theme_dag() 
mcar
```
\newpage

# MAR

( _ggdag_ draws bidrectional arrows for curved edges => need to fix manually)

```{r}
#  using more R-like syntax to create the same DAG
mar <- ggdag::dagify(
  Y ~ E,
  Y ~ c1 + C,
  E ~ c1 + C,
  C ~ c1,
  C_obs ~ M + C,
  M ~~ c1,
  exposure = "E",
  outcome = "Y",
  coords = coords)  %>% 
  ggdag::tidy_dagitty() %>% 
  ggdag::ggdag_classic() +
  ggdag::theme_dag() 
mar
```
\newpage

```{r}
# coordinates for mnar_u
coords_u <- list(
  x = c(c1 = -1, E = 0, Y = 1, C = -1, C_obs = -1, M = -1, U = 0),
  y = c(c1 =  2, E = 1, Y = 1, C =  0, C_obs = -1, M = -2, U = -1)
  )

#  using more R-like syntax to create the same DAG
mnar_u <- ggdag::dagify(
  Y ~ E,
  Y ~ c1 + C + U,
  E ~ c1 + C + U,
  C ~ c1,
  C_obs ~ M + C,
  M ~ U,
  exposure = "E",
  outcome = "Y",
  coords = coords_u)  %>% 
  ggdag::tidy_dagitty() %>% 
  ggdag::ggdag_classic() +
  ggdag::theme_dag() 
mnar_u
```

\newpage

# MNAR (value)

Missingness depends on the value of the confounder itself
( _ggdag_ draws bidrectional arrows for curved edges => need to fix manually)
```{r}
#  using more R-like syntax to create the same DAG
mnar_value <- ggdag::dagify(
  Y ~ E,
  Y ~ c1 + C,
  E ~ c1 + C,
  C ~ c1,
  C_obs ~ M + C,
  M ~~ C,
  exposure = "E",
  outcome = "Y",
  coords = coords)  %>% 
  ggdag::tidy_dagitty() %>% 
  ggdag::ggdag_classic() +
  ggdag::theme_dag() 
mnar_value
```
\newpage

# MNAR (unmeasured)

Missingness depends on unobserved factor.

\newpage

# Combined

Combine all plots in a panel

```{r, fig_width = 10, fig.height=10}
dag <- ggpubr::ggarrange(
  mcar, 
  mar, 
  mnar_u,
  mnar_value, 
  labels = c("a)", 
             "b)", 
             "c)", 
             "d)"),
  font.label = 
    list(
      size = 18, 
      color = "black", 
      face = "bold", 
      family = NULL),
  vjust = 1.25
  )
png(here::here("images", "dag_raw.png"), width = 12, height = 10, units = "in", res=300)
dag
invisible(dev.off())
```

```{r, fig_width = 10, fig.height=10}
dag
```

# Appendix

Scenario with c0

```{r}
# coordinates for mcar, mar, mnar_vaue
coords_c0 <- list(
  x = c(c0 = -2, c1 = -1, E = 0, Y = 1, C = -1, C_obs = -1, M = -1, U = -2),
  y = c(c0 = 1, c1 =  2, E = 1, Y = 1, C =  0, C_obs = -1, M = -2, U = -1)
  )
```

```{r}
#  using more R-like syntax to create the same DAG
# coordinates for mnar_u
coords_u <- list(
  x = c(c1 = -1, E = 0, Y = 1, C = -1, C_obs = -1, M = -1, U = 0, c0 = 1),
  y = c(c1 =  2, E = 1, Y = 1, C =  0, C_obs = -1, M = -2, U = -1, c0 = -2)
  )

#  using more R-like syntax to create the same DAG
mnar_u_c0 <- ggdag::dagify(
  Y ~ c1 + C + U,
  E ~ c1 + C + U,
  C ~ c1,
  C_obs ~ M + C,
  M ~ U,
  U ~ c0,
  exposure = "E",
  outcome = "Y",
  coords = coords_u)  %>% 
  ggdag::tidy_dagitty() %>% 
  ggdag::ggdag_classic() +
  ggdag::theme_dag() 
mnar_u_c0
```